A meta-analysis of fungicide spray schedules to minimise yield loss from powdery mildew in Australian mungbean

### 2021-02-11 01_TrialsForInclusionInAnalysis.Rmd
Rearrange code chunks to be more streamline and eliminate repetition

## Excluding variance
Following an inspection of the yield and disease variance we determined imputation was not suitable and therefore we must exlude trials that did not report variance.
For each analysis, yield and disease severity a separate data-set will be created, including only trials that report variance for the respective response variable

Four trials will be removed for our yield data set `PM_dat_Y` for not reporting yield variance.  
```{r TrialsWithVariance}
NoVar <- PM_MB_dat %>%
   mutate(No_Var = is.na(yield_error)) %>%
   filter(No_Var == TRUE) %>%
   distinct(trial_ref, year, location)
NoVar

PM_dat_Y <- PM_MB_dat %>%
   filter(!(trial_ref %in% NoVar$trial_ref))
```

Four trials will be removed for our disease severity data set `PM_dat_S` for not reporting yield variance.  
```{r TrialsWithVariance_D}
NoVar <- PM_MB_dat %>%
   mutate(No_Var = is.na(disease_error)) %>%
   filter(No_Var == TRUE) %>%
   distinct(trial_ref, year, location)
NoVar

PM_dat_D <- PM_MB_dat %>%
   filter(!(trial_ref %in% NoVar$trial_ref))
```

We will retain the following trials, which only contain DMI fungicides. 
However, we won't remove non-DMI fungicides just yet, we will do this before the analysis.

```{r Final_Trials_For_Inclusion}
PM_MB_dat %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   mutate(MA_analysis = case_when(
      is.na(yield_error & disease_error) == FALSE ~ "Yield and Disease",
      is.na(yield_error) == FALSE & 
         is.na(disease_error) ~ "Yield",
      is.na(disease_error) == FALSE & 
         is.na(yield_error) ~ "Disease",
      TRUE ~ "Nil"
   )) %>%
   distinct(trial_ref, location, year, MA_analysis) %>%
   arrange(year) 
```

<br>
<br>
<br>
<br>

@@

# classify yield data
PM_dat_Y <- 
   classify_PM_data(PM_dat_Y)

#classify disease data
PM_dat_D <- 
   classify_PM_data(PM_dat_D)
```

@@

```{r clustered_fungicide_applications}
# Yield data
PM_dat_Y %<>%
   mutate(fungicide_timing_1 = fungicide_application_1 - first_sign_disease) %>%
   mutate(fungicide_timing_2 = fungicide_application_2 - fungicide_application_1) %>%
   mutate(fungicide_timing_3 = fungicide_application_3 - fungicide_application_2)

# Disease data
PM_dat_D %<>%
   mutate(fungicide_timing_1 = fungicide_application_1 - first_sign_disease) %>%
   mutate(fungicide_timing_2 = fungicide_application_2 - fungicide_application_1) %>%
   mutate(fungicide_timing_3 = fungicide_application_3 - fungicide_application_2)

@@

Simplify the clusters for the yield data-set
```{r simple_Y_clusters}
PM_dat_Y <- PM_dat_Y %>%
   mutate(
      spray_management = case_when(
         fungicide_timing_1 < 0 &
            is.na(fungicide_application_2) &
            is.na(fungicide_application_3) ~ "Early",
         fungicide_timing_1 >= 0 &
            fungicide_timing_1 < 4 &
            is.na(fungicide_application_2) &
            is.na(fungicide_application_3) ~ "Recommended",
         fungicide_timing_1 >= 4 &
            is.na(fungicide_application_2) &
            is.na(fungicide_application_3) ~ "Late",
         fungicide_timing_1 < 0 &
            !is.na(fungicide_application_2) ~ "Early_plus",
         fungicide_timing_1 >= 0 &
            fungicide_timing_1 < 4 &
            !is.na(fungicide_application_2) ~ "Recommended_plus",
         fungicide_timing_1 >= 4 &
            !is.na(fungicide_application_2) ~ "Late_plus",
         TRUE ~ "Other"
      )
   )

PM_dat_Y[PM_dat_Y$fungicide_ai == "control",
          c(
             "fungicide_timing_1",
             "fungicide_timing_2",
             "fungicide_timing_3",
             "spray_management"
          )] <- "control"
```

Simplify the clusters
```{r simple_D_clusters}
PM_dat_D <- PM_dat_D %>%
   mutate(
      spray_management = case_when(
         fungicide_timing_1 < 0 &
            is.na(fungicide_application_2) &
            is.na(fungicide_application_3) ~ "Early",
         fungicide_timing_1 >= 0 &
            fungicide_timing_1 < 4 &
            is.na(fungicide_application_2) &
            is.na(fungicide_application_3) ~ "Recommended",
         fungicide_timing_1 >= 4 &
            is.na(fungicide_application_2) &
            is.na(fungicide_application_3) ~ "Late",
         fungicide_timing_1 < 0 &
            !is.na(fungicide_application_2) ~ "Early_plus",
         fungicide_timing_1 >= 0 &
            fungicide_timing_1 < 4 &
            !is.na(fungicide_application_2) ~ "Recommended_plus",
         fungicide_timing_1 >= 4 &
            !is.na(fungicide_application_2) ~ "Late_plus",
         TRUE ~ "Other"
      )
   )

PM_dat_D[PM_dat_D$fungicide_ai == "control",
          c(
             "fungicide_timing_1",
             "fungicide_timing_2",
             "fungicide_timing_3",
             "spray_management"
          )] <- "control"
```

@@

```{r tableSprayManagement_Y}
table(PM_dat_Y$spray_management)
```

```{r tableSprayManagement_Y}
table(PM_dat_D$spray_management)
```

@@

```{r simpler_clusters_remove}
PM_dat_Y <-
   PM_dat_Y %>%
   filter(spray_management != "Early_plus") %>%
   select(-c(fungicide_timing_1, # remove the following columns which no longer have use
             fungicide_timing_2,
             fungicide_timing_3,
             fungicide_application_1,
             fungicide_application_2,
             fungicide_application_3,
             fungicide_application_4,
             fungicide_application_5,
             fungicide_application_6,
             fungicide_application_7,
             ))

PM_dat_D <-
   PM_dat_D %>%
   filter(spray_management != "Early_plus") %>%
   select(-c(fungicide_timing_1, # remove the following columns which no longer have use
             fungicide_timing_2,
             fungicide_timing_3,
             fungicide_application_1,
             fungicide_application_2,
             fungicide_application_3,
             fungicide_application_4,
             fungicide_application_5,
             fungicide_application_6,
             fungicide_application_7,
             ))
```

>>>>

@@

# classify dataset with Trials in both 
PM_MB_dat <-
   classify_PM_data(PM_MB_dat)

@@

```{r clustered_fungicide_applications}
# collective data
PM_MB_dat %<>%
   mutate(fungicide_timing_1 = fungicide_application_1 - first_sign_disease) %>%
   mutate(fungicide_timing_2 = fungicide_application_2 - fungicide_application_1) %>%
   mutate(fungicide_timing_3 = fungicide_application_3 - fungicide_application_2)

@@

Simplify the clusters for the data-set
```{r simple_Y_clusters}
PM_MB_dat <- PM_MB_dat %>%
   mutate(
      spray_management = case_when(
         fungicide_timing_1 < 0 &
            is.na(fungicide_application_2) &
            is.na(fungicide_application_3) ~ "Early",
         fungicide_timing_1 >= 0 &
            fungicide_timing_1 < 4 &
            is.na(fungicide_application_2) &
            is.na(fungicide_application_3) ~ "Recommended",
         fungicide_timing_1 >= 4 &
            is.na(fungicide_application_2) &
            is.na(fungicide_application_3) ~ "Late",
         fungicide_timing_1 < 0 &
            !is.na(fungicide_application_2) ~ "Early_plus",
         fungicide_timing_1 >= 0 &
            fungicide_timing_1 < 4 &
            !is.na(fungicide_application_2) ~ "Recommended_plus",
         fungicide_timing_1 >= 4 &
            !is.na(fungicide_application_2) ~ "Late_plus",
         TRUE ~ "Other"
      )
   )

PM_MB_dat[PM_MB_dat$fungicide_ai == "control",
          c(
             "fungicide_timing_1",
             "fungicide_timing_2",
             "fungicide_timing_3",
             "spray_management"
          )] <- "control"
```

@@

```{r tableSprayManagement_Y}
table(PM_MB_dat$spray_management)
```

@@

```{r simpler_clusters_remove}
PM_MB_dat <-
   PM_MB_dat %>%
   filter(spray_management != "Early_plus") %>%
   select(-c(fungicide_timing_1, # remove the following columns which no longer have use
             fungicide_timing_2,
             fungicide_timing_3,
             fungicide_application_1,
             fungicide_application_2,
             fungicide_application_3,
             fungicide_application_4,
             fungicide_application_5,
             fungicide_application_6,
             fungicide_application_7,
             ))

## Trial variance
We need variance for each treatment for the data to be useful in a meta-analysis.
Lets look at which of the included trials contain variance for both response variables of interest, yield and disease severity.

```{r Trials_For_Inclusion}
PM_MB_dat %>%
   mutate(MA_analysis = case_when(
      is.na(yield_error & disease_error) == FALSE ~ "Yield and Disease",
      is.na(yield_error) == FALSE & 
         is.na(disease_error) ~ "Yield",
      is.na(disease_error) == FALSE & 
         is.na(yield_error) ~ "Disease",
      TRUE ~ "Nil"
   )) %>%
   distinct(trial_ref, location, year, MA_analysis) %>%
   arrange(year)

# Remove trials without variance for both Disease and yield for data_set `PM_MB_dat
PM_MB_dat <- 
   PM_MB_dat %>%
   mutate(contains_var = case_when(
      is.na(yield_error) &
         is.na(disease_error) ~ NA_character_,
      TRUE ~ trial_ref)) %>%
   filter(trial_ref %in% contains_var) %>%
   select(-contains_var)
```

## Excluding variance
Following an inspection of the yield and disease variance we determined imputation was not suitable and therefore we must exclude trials that did not report variance.
For each analysis, yield and disease severity a separate data-set will be created, including only trials that report variance for the respective response variable

Four trials will be removed for our yield data set `PM_dat_Y` for not reporting yield variance.  
```{r TrialsWithVariance}
PM_MB_dat %>%
   mutate(No_Var = is.na(yield_error)) %>%
   filter(No_Var == TRUE) %>%
   distinct(trial_ref, year, location)

PM_dat_Y <- 
   PM_MB_dat %>%
   mutate(contains_var = case_when(
      is.na(yield_error) ~ NA_character_,
      TRUE ~ trial_ref)) %>%
   filter(trial_ref %in% contains_var) %>%
   select(-contains_var)
```

Four trials will be removed for our disease severity data set `PM_dat_D` for not reporting yield variance.  
```{r TrialsWithVariance_D}
PM_MB_dat %>%
   mutate(No_Var = is.na(disease_error)) %>%
   filter(No_Var == TRUE) %>%
   distinct(trial_ref, year, location)

PM_dat_D <- 
   PM_MB_dat %>%
   mutate(contains_var = case_when(
      is.na(disease_error) ~ NA_character_,
      TRUE ~ trial_ref)) %>%
   filter(trial_ref %in% contains_var) %>%
   select(-contains_var)
```
